Online Learning of Per-flow Queuing Policies using Deep Reinforcement Learning

### 2020-06-28 paper/paper.tex
Paper draft finished

\begin{abstract}
The increasing number of different, incompatible congestion control algorithms has led to an increased deployment of fair queuing. Fair queuing isolates each network flow and gives it a separate queue. It can thus guarantee fairness for each flow even if the flows' congestion controls are not inherently fair. So far, each queue in the fair queuing system either has a fixed, static maximum size or is managed by an Active Queue Management algorithm (AQM) like CoDel. In this paper we design and implement an AQM mechanism that dynamically learns the optimal buffer size for each flow with respect to a specified utility function. We show that our Deep Learning based algorithm can correctly fingerprint each flow and assign the optimal queue size. Besides that we also show that our mechanism can even learn completely on-line and can continuously adapt. Finally we demonstrate that the computational overhead of our approach is sufficiently low to allow for deployment on regular hardware. 
\end{abstract}

@@

New \gls{cca} for TCP and QUIC are still being introduced, with one of the most prominent ones in recent years being BBR \cite{cardwell_bbr:_2016}\todo{Add more refs}. While new \glspl{cca} are commonly being designed with compatibility to other \glspl{cca} in mind, often they do not share the link completely fairly with older \glspl{cca}\todo{Add more refs}. Besides that, network flows using the same \gls{cca} can also be unfair to each other: For example, BBR favors flows with a high \gls{rtt}, while New Reno favors those with a low one \cite{turkovic_interactions_2019,turkovic_fifty_2019}. This unfairness can be mitigated by using \gls{fq} at the bottleneck link, isolating each flow from all other flows and assigning each flow an equal share of bandwidth \cite{dumazet_pkt_sched:_2013}.

The question of how to manage each queue then arises: The simplest solution is using a static buffer size (queue size) for each flow. Another solution is to use an advanced \gls{aqm} mechanism like CoDel for each flow. CoDel aims to keep the queue length under a certain threshold, making sure that the queuing delay was smaller than 5\,ms in the last 100\,ms at least once. Otherwise it drops packets to decrease the queue length. Both of these approaches do not differentiate between flows: They apply the same logic to each flow no matter its congestion control and no matter the current bandwidth and \gls{rtt}. \cite{bachl_cocoa_2019} showed that this behavior leads to some flows not being able to claim the full bandwidth that they are entitled to. Other flows might achieve the full bandwidth but keep an unnecessary standing queue. \cite{bachl_rax_2019} proposed a mechanism that adjusts the queue of each flow based on its congestion control and showed that it works well for several common \glspl{cca}. However, their algorithm has parameters that have to be manually adjusted and is not guaranteed to work for every \gls{cca} because it makes the assumption that each flow's congestion window follows a zigzag pattern. Furthermore, their approach is only tailored towards flows that always have data to send while the behavior for application limited flows, that also have idle periods, is not clear.

@@

@@

\begin{align}
\textit{Utility} = \textit{bandwidth}-\alpha\times\textit{queue size}
\end{align}

@@

@@

%\begin{table}
%\caption{Comparing the average throughput and maximum/average queue size over a large range of network conditions for New Reno and Bic. All results are the average of 100 experiments with delay ranging from 5 to 25\,ms (with bandwidth fixed at 15\,Mbit/s and of 100 experiments with bandwidth ranging from 5 to 25\,Mbit/s (with bandwidth fixed at 15\,ms). All experiments were conducted for both the New Reno and Bic \glspl{cca}. This gives a total of 400 experiments, over which the values are averaged.} \label{tab:comparison_others}
%\centering
%\begin{tabular}{r r r r r} \toprule
%& \gls{ours} & FqCoDel & Fifo 100 & Fifo 1000 \\
%\midrule
%avg.~throughput & 13.4 & \textbf{13.7} & 11.7 & 11.9 \\
%avg.~queue & \textbf{7.7} & 15.4 & 51.1 & 630.4 \\
%max.~queue & \textbf{23.9} & 155.4 & 100 & 1000 \\
%\bottomrule
%\end{tabular}
%\end{table}

>>>>

\begin{abstract}
The increasing number of different, incompatible congestion control algorithms has led to an increased deployment of fair queuing. Fair queuing isolates each network flow and gives it a separate queue. It can thus guarantee fairness for each flow even if the flows' congestion controls are not inherently fair. So far, each queue in the fair queuing system either has a fixed, static maximum size or is managed by an Active Queue Management algorithm (AQM) like CoDel. In this paper we design and implement an AQM mechanism that dynamically learns the optimal buffer size for each flow with respect to a specified utility function. We show that our Deep Learning based algorithm can correctly fingerprint each flow and assign the optimal queue size. Besides that we also show that our mechanism can even learn completely online and can continuously adapt. Finally, we show that the computational overhead of our approach is not excessive as it can be tuned according to requirements. 
\end{abstract}

@@

New \gls{cca} for TCP and QUIC are still being introduced revisiting old ideas but also introducing new concepts \cite{dong_pcc_2018, cardwell_bbr:_2016, hock_tcp_2017, bachl_cocoa_2019, jay_deep_2019}. BBR, for example, already being used in real-world production systems. While new \glspl{cca} are commonly being designed with compatibility to other \glspl{cca} in mind, often they do not share the link completely fairly with older \glspl{cca} \cite{hock_experimental_2017}. Besides that, network flows using the same \gls{cca} can also be unfair to each other: For example, BBR favors flows with a high \gls{rtt}, while New Reno favors those with a low one \cite{turkovic_interactions_2019,turkovic_fifty_2019}. This unfairness can be mitigated by using \gls{fq} at the bottleneck link, isolating each flow from all other flows and assigning each flow an equal share of bandwidth \cite{dumazet_pkt_sched:_2013}.

The question of how to manage each queue then arises: The simplest solution is using a static buffer size (queue size) for each flow. Another solution is to use an advanced \gls{aqm} mechanism like CoDel for each flow. CoDel aims to keep the queue length under a certain threshold, making sure that the queuing delay was smaller than 5\,ms in the last 100\,ms at least once. Otherwise it drops packets to decrease the queue length. Both of these approaches do not differentiate between flows: They apply the same logic to each flow no matter its congestion control and no matter the current bandwidth and \gls{rtt}. \cite{bachl_cocoa_2019} showed that this behavior leads to some flows not being able to claim the full bandwidth that they are entitled to. Other flows might achieve the full bandwidth but keep an unnecessary standing queue. \cite{bachl_cocoa_2019} proposed a mechanism that adjusts the queue of each flow based on its congestion control and showed that it works well for several common \glspl{cca}. However, their algorithm has parameters that have to be manually adjusted and is not guaranteed to work for every \gls{cca} because it makes the assumption that each flow's congestion window follows a zigzag pattern. Furthermore, their approach is only tailored towards flows that always have data to send while the behavior for application limited flows, that also have idle periods, is not clear.

@@

\section{Related Work}

\cite{bachl_cocoa_2019} uses a hand-crafted algorithm to optimize the queuing behavior for different \glspl{cca}. It is, however, not guaranteed that this algorithm works for all new \glspl{cca} and for all traffic pattern.

\cite{kim_deep_2019} use Deep Reinforcement Learning to develop an \gls{aqm} mechanism for IoT devices that works on the output queue of a switch.  \cite{bouacida_practical_2019,bisoy_design_2017} use \gls{rl} to build an \gls{aqm} algorithm. \cite{vucevic_reinforcement_2007} uses \gls{rl} as well and focuses specifically on cellular communications. \cite{lin_kemy_2015} use a custom optimization approach that relies on as offline greedy algorithm which finds optimal rules for an \gls{aqm} that fulfills a certain utility function. \cite{shah_sam_2016} build an \gls{aqm} mechanism, which utilizes support vector machines to learn an optimal queuing policy for a shared buffer.

Unlike this work, these works' aim is not fingerprinting each flow but instead they optimize the overall queue that is shared by flows. In our work we focus on fingerprinting each flow and providing the optimal queuing behavior for each flow individually. This direction has, to our knowledge, not been explored extensively.

@@

\begin{align}
\textit{Utility} = \textit{bandwidth}-\alpha\times\textit{queue size}
\label{eq:utility}
\end{align}

@@

Comparing the behavior of \gls{ours} (\autoref{fig:exampleBicOursNice}) with the behavior of Fifo (\autoref{fig:exampleBicFifo1000} and \autoref{fig:exampleBicFifo100}) for an example flow shows that while Fifo keeps a very large standing queue, \gls{ours} keeps the queue at the minimum that is possible to achieve maximum throughput and minimum queue according to our utility function (\autoref{eq:utility}). The comparison with FqCoDel (\autoref{fig:exampleBicFqCoDel}) shows that FqCoDel generally avoids a large standing queue but results in a very large spike in queued packets at the beginning of the flow. Thus, while CoDel was designed to avoid ``bufferbloat'' it actually induces a very large standing queue during the slow start phase of a flow.

@@

The systematic comparison (\autoref{tab:comparison_others}) between the different fair \gls{aqm} mechanism shows that \gls{ours} achieves about the same or better throughput than the competing algorithms while having a significantly smaller average queue and an even more pronounced reduction in maximum queue when compared to Fifo and FqCoDel.  
%\begin{table}
%\caption{Comparing the average throughput and maximum/average queue size over a large range of network conditions for New Reno and Bic. All results are the average of 100 experiments with delay ranging from 5 to 25\,ms (with bandwidth fixed at 15\,Mbit/s and of 100 experiments with bandwidth ranging from 5 to 25\,Mbit/s (with bandwidth fixed at 15\,ms). All experiments were conducted for both the New Reno and Bic \glspl{cca}. This gives a total of 400 experiments, over which the values are averaged.} \label{tab:comparison_others}
%\centering
%\begin{tabular}{r r r r r} \toprule
%& \gls{ours} & FqCoDel & Fifo 100 & Fifo 1000 \\
%\midrule
%avg.~throughput & 13.4 & \textbf{13.7} & 11.7 & 11.9 \\
%avg.~queue & \textbf{7.7} & 15.4 & 51.1 & 630.4 \\
%max.~queue & \textbf{23.9} & 155.4 & 100 & 1000 \\
%\bottomrule
%\end{tabular}
%\end{table}
